---
title: "Clinical Statistics Methods Forum Data Challenge"
author: "Eric Polley & Prabin Thapa"
date: "October 18th, 2016"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

## Introduction

The 2016 Clinical Statistics Methods Forum Data Challenge

The goal is to construct a predictor of therapeutic drug dose given baseline clinical measurements. Participants given a training data set with 3,000 patients and will be evaluated on a blinded validation data set with 1,722 patients

## FAQ

- Teams can submit as many predictions on the validation data set as they want, but I'll randomly select only one for evaluation
- Two primary loss functions:
\begin{equation}
L_{A}(\hat{y}, y) = (\hat{y} - y)^2
\end{equation}
And
\begin{equation}
L_{B}(\hat{y}, y) = \mathrm{I}\!\left(\lvert \hat{y}-y\rvert > 16\right)
\end{equation}
- The equation $L_B(x, y)$ is the indicator the predicted dose is more than 1 standard deviations away from the truth
- These notes will be available at \url{https://github.com/ecpolley/CSMF_Data_Challenge}


## Plan for November

- Please submit predictions on the validation data prior to\newline Nov. 14th, 5:00PM (central)
- Meet on Nov. 15th, 2:00PM
- Each team will have the opportunity to describe their analytic approach with 1-2 slides
- Briefly describe how you processed the data and your approach for estimating the predictor
- I will then summarize the different methods and reveal the validation set performance with discussion 

## How To Get Data

```{r loadData, echo = TRUE}
# link to data on GitHub page if not available
if(file.exists("TRAIN.CSV")) {
  TRAIN <- read.csv("TRAIN.CSV")
} else {
  urlfile <- "https://raw.githubusercontent.com/ecpolley/
    CSMF_Data_Challenge/master/TRAIN.CSV"
  download.file(urlfile, destfile = "TRAIN.CSV")
  TRAIN <- read.csv("TRAIN.CSV")
}
dim(TRAIN)
```

## How To Get Data

```{r loadData2, echo = TRUE}
# link to data on GitHub page if not available
if(file.exists("VALID.CSV")) {
  VALID <- read.csv("VALID.CSV")
} else {
  urlfile <- "https://raw.githubusercontent.com/ecpolley/
    CSMF_Data_Challenge/master/VALID.CSV"
  download.file(urlfile, destfile = "VALID.CSV")
  VALID <- read.csv("VALID.CSV")
}
dim(VALID)
```

## Biomarkers

- Show R code

## Data processing

- How are teams working with the variables?
- Did you create any new variables?
- What about missing values?

## Methods 

- Which prediction methods have you considered?

## Loss functions

- Have you considered different loss functions other than squared error?
- What are you using to evaluate the performance of a predictor?


## Beyond Analytics

- If you are working in a team with multiple individuals, how are you dividing the project?

## Questions?