---
title: "Clinical Statistics Methods Forum Data Challenge"
author: "Eric Polley & Prabin Thapa"
date: "September 20th, 2016"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

## Introduction

The 2016 Clinical Statistics Methods Forum Data Challenge

The goal is to construct a predictor of therapeutic drug dose given baseline clinical measurements. Participants given a training data set with 3,000 patients and will be evaluated on a blinded validation data set with 1,722 patients

## Overview

- Working in teams, n $\in$ (1, 2, \ldots, 20), use the training data to construct a predictor for the therapeutic dose given baseline clinical measurements
- Predictors evaluated by mean squared error of predicted dose and true dose
- With validation data set, provide predictions for each individual. A file with the Subject ID and the predicted dose as 2 columns can be email to me
- Don't forget to include your team name
- On Oct.\ 18th, meet for a midpoint review and group discussion
- On Nov.\ 15th, each team will be asked to provide a short summary on how they constructed the predictor and I will reveal the  performance on the validation data
- Everyone is welcome to attend the discussion (and drink coffee)
- Prizes to be determined

## Overview

- 2 data sets provided in comma separated format
- TRAIN.CSV includes the baseline variables and outcome of interest (Therapeutic_Dose)
- VALID.CSV includes the baseline variables, need to predict the outcome
- Data available on GitHub: \url{https://github.com/ecpolley/CSMF_Data_Challenge}

## How To Get Data

```{r loadData, echo = TRUE}
# link to data on GitHub page if not available
if(file.exists("TRAIN.CSV")) {
  TRAIN <- read.csv("TRAIN.CSV")
} else {
  urlfile <- "https://raw.githubusercontent.com/ecpolley/
    CSMF_Data_Challenge/master/TRAIN.CSV"
  download.file(urlfile, destfile = "TRAIN.CSV")
  TRAIN <- read.csv("TRAIN.CSV")
}
dim(TRAIN)
```

## Therapeutic_Dose

```{r histogram}
ggplot(TRAIN, aes(Therapeutic_Dose)) + geom_histogram(bins = 30) + theme(text = element_text(size=30))
```

## Baseline Variables
```{r echo=FALSE}
knitr::knit_hooks$set(mysize = function(before, options, envir) {
  if (before) 
    return(options$size)
})
```

```{r str1, mysize = TRUE, size = '\\footnotesize', echo = TRUE}
str(TRAIN[, 1:8])
```

## Baseline Variables
```{r plotAge, out.width='5.5cm', out.height='7cm', fig.show='hold'}
ggplot(TRAIN, aes(x = Age, y = Therapeutic_Dose)) + geom_boxplot() + theme(text = element_text(size=30))
ggplot(TRAIN, aes(x = Gender, y = Therapeutic_Dose)) + geom_boxplot() + theme(text = element_text(size=30))
```


## Baseline Variables

```{r str2, mysize = TRUE, size = '\\footnotesize', echo = TRUE}
str(TRAIN[, 9:12])
```

## Baseline Variables

```{r str3, mysize = TRUE, size = '\\footnotesize', echo = TRUE}
str(TRAIN[, 13:29])
```

## Baseline Variables

```{r str4, mysize = TRUE, size = '\\footnotesize', echo = TRUE}
str(TRAIN[, 30:33])
```

## Baseline Variables

```{r str5, mysize = TRUE, size = '\\footnotesize', echo = TRUE}
str(TRAIN[, 34:43])
```

## Basic Regression Predictor

```{r lmFit, mysize = TRUE, size = '\\footnotesize', echo = TRUE}
fit <- lm(Therapeutic_Dose ~ Age + Gender + 
            Biomarker_1 + Biomarker_2, data = TRAIN)
fit
```


<!-- ## Basic Regression Predictor -->

<!-- ```{r lmFit3, mysize = TRUE, size = '\\footnotesize', echo = TRUE} -->
<!-- mean((predict(fit) - TRAIN$Therapeutic_Dose)^2) -->
<!-- ``` -->

## How To Get Data

```{r loadData2, echo = TRUE}
# link to data on GitHub page if not available
if(file.exists("VALID.CSV")) {
  VALID <- read.csv("VALID.CSV")
} else {
  urlfile <- "https://raw.githubusercontent.com/ecpolley/
    CSMF_Data_Challenge/master/VALID.CSV"
  download.file(urlfile, destfile = "VALID.CSV")
  VALID <- read.csv("VALID.CSV")
}
dim(VALID)
```

## Basic Regression Predictor

```{r lmPred, echo = TRUE}
pred_lm <- predict(fit, newdata = VALID)
pred_lm_df <- data.frame(ID = VALID$Subject_ID, predict = pred_lm)
head(pred_lm_df)
# write out prediction table
# write.csv(pred_lm_df, "final_predictions.csv")  
```

## Cross-Validation
Prior to submitting predictions, should evaluate performance
```{r cv, echo = TRUE}
## get V-fold CV estimate of MSE
V <- 10
N <- nrow(TRAIN)
MSE_cv <- rep(NA, V)  # placeholder for CV MSE estimates
# list of row ids by V validation splits
validRows <- split(sample(1:N), rep(1:V, length=N)) 
for(v in seq(V)) {
  tempTRAIN <- TRAIN[-validRows[[v]], ]
  tempVALID <- TRAIN[validRows[[v]], ]
  fit_cv <- lm(Therapeutic_Dose ~ Age + Gender + 
            Biomarker_1 + Biomarker_2, data = tempTRAIN)
  pred_cv <- predict(fit_cv, newdata = tempVALID)
  MSE_cv[v] <- mean((pred_cv - tempVALID$Therapeutic_Dose)^2)
}
```

## Cross-Validation
```{r cv_output, echo = TRUE}
# get V-fold CV estimate of MSE
summary(MSE_cv)
# CV R-squared
1 - mean(MSE_cv)/var(TRAIN$Therapeutic_Dose)
```

## Questions?